---
title: knb-lter-cap.41
author: CAP LTER
---

## pitfalls: query

*careful to omit `traps.trap_name::INTEGER` for core sites as some have non-numeric trap names*

```{r}
#| label: pitfalls-query
#| eval: true

mcdowell_pitfall_arthropods <- DBI::dbGetQuery(
  conn      = pg,
  statement = "
  WITH trap_counts AS (
    SELECT
      trap_sampling_events.sampling_event_id,
      COUNT(DISTINCT trap_sampling_events.id) AS trap_count
    FROM arthropods.trap_sampling_events
    WHERE trap_sampling_events.flags IS DISTINCT FROM 'trap_not_collected'
    GROUP BY trap_sampling_events.sampling_event_id
  )
  SELECT
    sites.site_code,
    sampling_events.sample_date,
    trap_sampling_events.comments,
    trap_sampling_events.flags,
    arthropod_taxonomy.display_name,
    trap_specimens.\"count\",
    trap_specimens.immature,
    trap_specimens.winged,
    size_classes.description AS size,
    observers.observer_name AS observer,
    traps.trap_name,
    trap_counts.trap_count
  FROM arthropods.sampling_events
  JOIN arthropods.sites
    ON arthropods.sampling_events.site_id = arthropods.sites.id
  JOIN arthropods.trap_sampling_events
    ON arthropods.trap_sampling_events.sampling_event_id = arthropods.sampling_events.id
  LEFT JOIN arthropods.trap_specimens
    ON arthropods.trap_specimens.trap_sampling_event_id = arthropods.trap_sampling_events.id
  LEFT JOIN arthropods.size_classes
    ON arthropods.trap_specimens.size_class_id = arthropods.size_classes.id
  LEFT JOIN arthropods.observers
    ON arthropods.observers.id = arthropods.trap_specimens.person_id
  LEFT JOIN arthropods.arthropod_taxonomy
    ON arthropods.arthropod_taxonomy.id = arthropods.trap_specimens.arthropod_taxon_id
  LEFT JOIN arthropods.traps
    ON arthropods.traps.id = arthropods.trap_sampling_events.trap_id
  LEFT JOIN trap_counts
  ON trap_counts.sampling_event_id = trap_sampling_events.sampling_event_id
  WHERE arthropods.sites.site_code IN (
    'TomThumb',
    'Gateway',
    'Bell',
    'Sunrise',
    'LoneMtn',
    'Paraiso',
    'Dixileta',
    'Rincon',
    'Mine',
    'Prospector'
  )
  ORDER BY
    sites.site_code,
    sampling_events.sample_date,
    arthropod_taxonomy.display_name
  ;
  "
)

```

Convert observer names to first two letters of each name, including a middle
name if provided. This approach provides anonymity yet distinctly identifies
each observer if needed as a covariate. Pulling this code out separately owing
to its verbosity for a singular purpose: getting the first two letters of the
first and last names of each observers, and presenting those instead of the full
name.

## pitfalls: observers

```{r}
#| eval: TRUE
#| label: pitfalls-observers

mcdowell_pitfall_arthropods <- mcdowell_pitfall_arthropods |>
  tidyr::separate_wider_delim(
    cols        = observer,
    names       = c("name1", "name2"),
    delim       = " ",
    cols_remove = FALSE
    ) |>
  dplyr::mutate(
    namePart1 = tools::toTitleCase(stringr::str_extract(name1, "\\b\\w{2}")),
    namePart2 = tools::toTitleCase(stringr::str_extract(name2, "\\b\\w{2}")),
    observer  = dplyr::if_else(
      is.na(namePart1) | is.na(namePart2),
      NA_character_,
      paste0(namePart1, namePart2)
    )
  ) |>
  dplyr::select(
    -name1,
    -name2,
    -tidyselect::contains("namePart")
  )

```

## pitfalls: conversions

```{r}
#| eval: TRUE
#| label: pitfalls-conversions

mcdowell_pitfall_arthropods <- mcdowell_pitfall_arthropods |>
  dplyr::mutate(
    site_code = as.factor(site_code),
    flags     = as.factor(flags),
    immature  = as.factor(immature),
    winged    = as.factor(winged),
    size      = as.factor(size)
  )

```

## pitfalls: dataTable

There are few numeric variables that are likely to change (meaningfully) so
`update_attributes` should generally not be required.

```{r}
#| eval: TRUE
#| label: pitfalls-DT

# capeml::update_attributes(mcdowell_pitfall_arthropods)

try({
  capeml::write_attributes(mcdowell_pitfall_arthropods, overwrite = FALSE)
  capeml::write_factors(mcdowell_pitfall_arthropods, overwrite = FALSE)
})

```

## sites: core sites metadata

```{r}
#| label: core-arthropod-sites
#| eval: TRUE

mcdowell_arthropod_sites <- DBI::dbGetQuery(
  conn      = pg,
  statement = "
  SELECT
    site_code,
    site_type,
    num_traps,
    trap_arrange,
    start_date,
    end_date,
    comments
  FROM arthropods.sites
    WHERE sites.site_code IN (
    'TomThumb',
    'Gateway',
    'Bell',
    'Sunrise',
    'LoneMtn',
    'Paraiso',
    'Dixileta',
    'Rincon',
    'Mine',
    'Prospector'
  )
  ;
  "
)

mcdowell_arthropod_sites <- mcdowell_arthropod_sites |>
  dplyr::mutate(site_type = as.factor(site_type))

try({
  capeml::write_attributes(mcdowell_arthropod_sites, overwrite = FALSE)
  capeml::write_factors(mcdowell_arthropod_sites, overwrite = FALSE)
})

```

## sites: core sites locations

As it is undesirable to acknowledge the precise location of the pitfall traps in
the McDowell Sonoran Preserve, the approach taken here is to create bounding
boxes around each of the paired sites: Brown's Ranch, Tom's Thumb, Gateway,
Dixie Mine, and Lost Dog. These spatial data are reflected in the file
`core_arthropod_locations.geojson`, which can be reconstructed from
`mcdowell_sites_polygons.R` if needed.

```{r}
#| label: core-arthropod-locations
#| eval: TRUE

core_arthropod_locations <- sf::st_read("core_arthropod_locations.geojson")

try({capeml::write_attributes(core_arthropod_locations, overwrite = FALSE)})

core_arthropod_locations_desc <- "Geospatial file detailing the locations of ground-dwelling arthropod sampling sites in the McDowell Sonoran Preserve, Scottsdale, Arizona. Polygons reflect the bounding box (minimum and maximum extents) of paired (boundary-interior) sampling locations. Precise sampling locations are available upon request to the CAP LTER Information Manager."

core_arthropod_locations_SV <- capemlGIS::create_vector(
  vector_name = core_arthropod_locations,
  description = core_arthropod_locations_desc,
  driver      = "GeoJSON"
)

```


## coverages

```{r}
#| eval: TRUE
#| label: coverages

coverage <- EML::set_coverage(
  begin                 = as.character(as.Date(min(mcdowell_pitfall_arthropods$sample_date))),
  end                   = as.character(as.Date(max(mcdowell_pitfall_arthropods$sample_date))),
  geographicDescription = capeml::read_package_configuration()[["geographic_description"]],
  west                  = as.numeric(sf::st_bbox(core_arthropod_locations)$xmin),
  east                  = as.numeric(sf::st_bbox(core_arthropod_locations)$xmax),
  north                 = as.numeric(sf::st_bbox(core_arthropod_locations)$ymax),
  south                 = as.numeric(sf::st_bbox(core_arthropod_locations)$ymin)
)

```

## taxonomic coverage

```{r}
#| label: taxonomyCleanr
#| eval: TRUE

my_path <- getwd()

taxa <- dplyr::bind_rows(
  mcdowell_pitfall_arthropods |>
    dplyr::select(display_name),
  mcdowell_vegetation_arthropods |>
    dplyr::select(display_name),
  mcdowell_vegetation_arthropods |>
    dplyr::select(display_name = plant_scientific_name)
) |>
  dplyr::mutate(display_name = gsub("^([A-Za-z]+) \\(.*\\)$", "\\1", display_name)) |>
  dplyr::distinct(display_name) |>
  as.data.frame()

taxonomyCleanr::create_taxa_map(
  path = my_path,
  x    = taxa,
  col  = "display_name"
)

taxonomyCleanr::resolve_sci_taxa(
  path         = my_path,
  data.sources = c(3, 11)
)

taxaCoverage <- taxonomyCleanr::make_taxonomicCoverage(path = my_path)

coverage$taxonomicCoverage <- taxaCoverage

```

```{r}
#| eval: TRUE
#| label: build

dataset <- capeml::create_dataset()
eml     <- capeml::create_eml()

EML::eml_validate(eml)
capeml::write_cap_eml()

source("~/Documents/localSettings/aws.s3")

lapply(
  X   = list.files(pattern = "41_"),
  FUN = capeml::data_to_amz
)

source("~/Documents/localSettings/edi.R")
capeml::get_package_evaluation(full_report = FALSE)

report <- capeml::get_package_evaluation(full_report = TRUE)
xml2::write_xml(report, "/tmp/report.xml")

capeml::create_package(environment = "production")

EDIutils::logout()
```

```{r}
#| eval: TRUE
#| label: clean

file.remove(c(list.files(pattern = "^41.+csv$")))
capeml::eml_to_amz(list.files(pattern = "knb.+xml"))
file.copy(list.files(pattern = "knb.+xml"), "/home/srearl/localRepos/cap-metadata/cap-data-eml/")
```
