---
title: knb-lter-cap.41
author: CAP LTER
---

## identify McDowell sites

```{r}
#| label: identify-mcdowell-sites 
#| eval: TRUE

mcdowell_sites <- c(
  "TomThumb",
  "Gateway",
  "Bell",
  "Sunrise",
  "LoneMtn",
  "Paraiso",
  "Dixileta",
  "Rincon",
  "Mine",
  "Prospector"
)

```


## pitfalls: query

*careful to avoid `traps.trap_name::INTEGER` for core sites as some have non-numeric trap names*

```{r}
#| label: pitfalls-query
#| eval: true

core_arthropods_query <- glue::glue_sql("
  WITH trap_counts AS (
    SELECT
      trap_sampling_events.sampling_event_id,
      COUNT(DISTINCT trap_sampling_events.id) AS trap_count
    FROM arthropods.trap_sampling_events
    WHERE trap_sampling_events.flags IS DISTINCT FROM 'trap_not_collected'
    GROUP BY trap_sampling_events.sampling_event_id
  )
  SELECT
    sites.site_code,
    sampling_events.sample_date,
    trap_sampling_events.comments,
    trap_sampling_events.flags,
    arthropod_taxonomy.display_name,
    arthropod_taxonomy.authority,
    arthropod_taxonomy.authority_id,
    trap_specimens.\"count\",
    trap_specimens.immature,
    trap_specimens.winged,
    size_classes.description AS size,
    observers.observer_name AS observer,
    traps.trap_name,
    trap_counts.trap_count
  FROM arthropods.sampling_events
  JOIN arthropods.sites
    ON arthropods.sampling_events.site_id = arthropods.sites.id
  JOIN arthropods.trap_sampling_events
    ON arthropods.trap_sampling_events.sampling_event_id = arthropods.sampling_events.id
  LEFT JOIN arthropods.trap_specimens
    ON arthropods.trap_specimens.trap_sampling_event_id = arthropods.trap_sampling_events.id
  LEFT JOIN arthropods.size_classes
    ON arthropods.trap_specimens.size_class_id = arthropods.size_classes.id
  LEFT JOIN arthropods.observers
    ON arthropods.observers.id = arthropods.trap_specimens.person_id
  LEFT JOIN arthropods.arthropod_taxonomy
    ON arthropods.arthropod_taxonomy.id = arthropods.trap_specimens.arthropod_taxon_id
  LEFT JOIN arthropods.traps
    ON arthropods.traps.id = arthropods.trap_sampling_events.trap_id
  LEFT JOIN trap_counts
  ON trap_counts.sampling_event_id = trap_sampling_events.sampling_event_id
  WHERE sites.site_code NOT IN ({ mcdowell_sites* })
  ORDER BY
    sites.site_code,
    sampling_events.sample_date,
    arthropod_taxonomy.display_name
  ;
  ",
  .con = ANSI()
)

core_arthropods <- DBI::dbGetQuery(
  conn      = pg,
  statement = core_arthropods_query 
)

```

Convert observer names to first two letters of each name, including a middle
name if provided. This approach provides anonymity yet distinctly identifies
each observer if needed as a covariate. Pulling this code out separately owing
to its verbosity for a singular purpose: getting the first two letters of the
first and last names of each observers, and presenting those instead of the full
name.

## pitfalls: observers

```{r}
#| eval: TRUE
#| label: pitfalls-observers

derive_observer_id <- function(name) {
  if (base::is.na(name) || name == "" || base::tolower(name) == "na" || stringr::str_to_lower(name) == "unknown") {
    return(NA_character_)
  }
  name <- stringr::str_trim(name)
  # preserve "Ceh" and "MarissaB" (exact match)
  if (name %in% c("Ceh", "MarissaB")) {
    return(name)
  }
  if (stringr::str_detect(name, "^[A-Z]{2,4}$")) {
    return(name)
  }
  parts <- stringr::str_extract_all(name, "\\b\\w+\\b")[[1]]
  if (base::length(parts) >= 2) {
    return(
      base::paste0(
        stringr::str_to_upper(stringr::str_sub(parts[1], 1, 2)),
        stringr::str_to_upper(stringr::str_sub(parts[base::length(parts)], 1, 2))
      )
    )
  } else if (base::length(parts) == 1) {
    return(stringr::str_to_upper(stringr::str_sub(parts[1], 1, 2)))
  } else {
    return(NA_character_)
  }
}

core_arthropods <- core_arthropods |>
  dplyr::mutate(observer = purrr::map_chr(
    .x = core_arthropods$observer,
    .f = derive_observer_id
  ))

```

## pitfalls: conversions

```{r}
#| eval: TRUE
#| label: pitfalls-conversions

core_arthropods <- core_arthropods |>
  dplyr::mutate(
    site_code    = as.factor(site_code),
    flags        = as.factor(flags),
    immature     = as.factor(immature),
    winged       = as.factor(winged),
    size         = as.factor(size),
    authority    = as.factor(authority),
    authority_id = as.character(authority_id)
  )

```

## pitfalls: dataTable

There are few numeric variables that are likely to change (meaningfully) so
`update_attributes` should generally not be required.

```{r}
#| eval: TRUE
#| label: pitfalls-DT

# capeml::update_attributes(core_arthropods)

try({
  capeml::write_attributes(core_arthropods, overwrite = FALSE)
  capeml::write_factors(core_arthropods, overwrite = FALSE)
})

```

## sites: core sites metadata

```{r}
#| label: core-arthropod-sites
#| eval: TRUE

core_arthropods_sites_query <- glue::glue_sql(
  conn      = pg,
  statement = "
  SELECT
    site_code,
    num_traps,
    trap_arrange,
    start_date,
    end_date,
    comments
  FROM arthropods.sites
  WHERE sites.site_code NOT IN ({ mcdowell_sites* })
  ;
  ",
  .con = ANSI()
)

core_arthropods_sites <- DBI::dbGetQuery(
  conn      = pg,
  statement = core_arthropods_sites_query 
) |>
  dplyr::mutate(site_type = as.factor(site_type))

try({capeml::write_attributes(core_arthropods_sites, overwrite = FALSE)})

```

## sites: core sites locations

```{r}
#| label: core-arthropod-locations
#| eval: TRUE

core_locations_query <- glue::glue_sql("
  SELECT
    site_code,
    lat,
    long
  FROM arthropods.sites
  WHERE sites.site_code NOT IN ({ mcdowell_sites* })
  ",
  .con = ANSI()
)

core_arthropods_locations <- DBI::dbGetQuery(
  conn      = pg,
  statement = core_locations_query
) |>
  sf::st_as_sf(
    coords = c("long", "lat"),
    crs    = 4326
  ) |>
  sf::st_transform(crs = 32612)

# function to create a square buffer (as an sfg polygon) around a point
generate_buffer <- function(geom, dist) {

  # extract coordinates from the point geometry
  coords <- sf::st_coordinates(geom)
  x <- coords[1]
  y <- coords[2]
  
  # define the square corners (clockwise or counterclockwise)
  square_coords <- matrix(
    c(
      x - dist, y - dist,
      x - dist, y + dist,
      x + dist, y + dist,
      x + dist, y - dist,
      x - dist, y - dist  # close the polygon
    ),
    ncol = 2,
    byrow = TRUE
  )
  
  # create an sfg polygon object
  square <- sf::st_polygon(list(square_coords))
  return(square)

}

site_buffers <- sf::st_sfc(
  purrr::map(sf::st_geometry(core_arthropods_locations), generate_buffer, dist = 50),
  crs = sf::st_crs(core_arthropods_locations)
)

core_arthropods_locations$geometry <- site_buffers

core_arthropods_locations <- core_arthropods_locations |>
  sf::st_transform(crs = 4326)

try({capeml::write_attributes(core_arthropods_locations, overwrite = FALSE)})

core_arthropods_locations_desc <- "Geospatial file detailing the locations of ground-dwelling arthropod sampling sites in the CAP LTER study area encompassing the greater Phoenix metropolitan area and immediately surrounding Sonoran desert region. Polygons reflect the approximate (within 50 meters) location of sampling sites. Precise sampling locations are available upon request to the CAP LTER Information Manager."

core_arthropods_locations_SV <- capemlGIS::create_vector(
  vector_name = core_arthropods_locations,
  description = core_arthropods_locations_desc,
  driver      = "GeoJSON"
)

```


## coverages

```{r}
#| eval: TRUE
#| label: coverages

coverage <- EML::set_coverage(
  begin                 = as.character(as.Date(min(core_arthropods$sample_date))),
  end                   = as.character(as.Date(max(core_arthropods$sample_date))),
  geographicDescription = capeml::read_package_configuration()[["geographic_description"]],
  west                  = as.numeric(sf::st_bbox(core_arthropods_locations)$xmin),
  east                  = as.numeric(sf::st_bbox(core_arthropods_locations)$xmax),
  north                 = as.numeric(sf::st_bbox(core_arthropods_locations)$ymax),
  south                 = as.numeric(sf::st_bbox(core_arthropods_locations)$ymin)
)

```

## taxonomic coverage

```{r}
#| label: taxonomyCleanr
#| eval: TRUE

my_path <- getwd()

taxa <- core_arthropods |>
  dplyr::mutate(display_name = gsub("^([A-Za-z]+) \\(.*\\)$", "\\1", display_name)) |>
  dplyr::distinct(display_name) |>
  as.data.frame()

taxonomyCleanr::create_taxa_map(
  path = my_path,
  x    = taxa,
  col  = "display_name"
)

taxonomyCleanr::resolve_sci_taxa(
  path         = my_path,
  data.sources = c(3, 11)
)

taxaCoverage <- taxonomyCleanr::make_taxonomicCoverage(path = my_path)

coverage$taxonomicCoverage <- taxaCoverage

```

```{r}
#| eval: TRUE
#| label: build

dataset <- capeml::create_dataset()
eml     <- capeml::create_eml()

EML::eml_validate(eml)
capeml::write_cap_eml()

source("~/Documents/localSettings/aws.s3")

lapply(
  X   = list.files(pattern = "41_"),
  FUN = capeml::data_to_amz
)

source("~/Documents/localSettings/edi.R")
capeml::get_package_evaluation(full_report = FALSE)

report <- capeml::get_package_evaluation(full_report = TRUE)
xml2::write_xml(report, "/tmp/report.xml")

capeml::create_package(environment = "development")

EDIutils::logout()

```

```{r}
#| eval: TRUE
#| label: clean

file.remove(c(list.files(pattern = "^41.+csv$")))
capeml::eml_to_amz(list.files(pattern = "knb.+xml"))
file.copy(list.files(pattern = "knb.+xml"), "/home/srearl/localRepos/cap-metadata/cap-data-eml/")

```
