---
title: "knb-lter-cap.41"
author: "SRE"
date: "March 14, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

# README

contents moved to README.md

# libraries

```{r libraries}
library(EML)
library(RPostgreSQL)
library(RMySQL)
library(tidyverse)
library(tools)
library(readxl)
library(aws.s3)
library(capeml)
library(gioseml)
```

# core arthropods

```{r core-arthropods}

core_arthropods <- dbGetQuery(mysql, "
SELECT  
  se.sampling_event_id,
  s.site_code,
  se.sample_date,
  coalesce(specimens_data.observer, outer_join_people.observer) AS observer,
  t.trap_name,
  tse.comments AS trap_sampling_events_comments,
  tse.flags AS trap_sampling_events_flags,
  count_data.trap_count,
  specimens_data.arth_class,
  specimens_data.arth_order,
  specimens_data.arth_family,
  specimens_data.arth_genus_subgenus,
  specimens_data.display_name,
  specimens_data.lt2mm,
  specimens_data._2_5mm,
  specimens_data._5_10mm,
  specimens_data.gt10mm,
  specimens_data.unsized
FROM lter10_arthropods_production.trap_sampling_events tse
JOIN lter10_arthropods_production.sampling_events se ON (se.sampling_event_id = tse.sampling_event_id)
JOIN lter10_arthropods_production.sites s ON (s.site_id = se.site_id)
JOIN lter10_arthropods_production.traps t ON (tse.trap_id = t.trap_id)
LEFT JOIN lter10_arthropods_production.people outer_join_people ON (outer_join_people.person_id = se.default_person_for_trap_samples)
LEFT JOIN lter10_arthropods_production.specimens_data ON (specimens_data.trap_sampling_event_id = tse.trap_sampling_event_id)
JOIN
(
  SELECT
  trap_sampling_events.sampling_event_id,
  COUNT(DISTINCT trap_sampling_events.trap_sampling_event_id) AS trap_count
  FROM lter10_arthropods_production.trap_sampling_events
  WHERE NOT 
  (
    trap_sampling_events.flags LIKE 'trap not collected' OR 
    trap_sampling_events.flags LIKE 'NotCollected' OR
    trap_sampling_events.flags LIKE 'missing'
  )  OR 
    trap_sampling_events.flags IS NULL OR
    trap_sampling_events.flags = ''
  GROUP BY sampling_event_id
) AS count_data ON (count_data.sampling_event_id = tse.sampling_event_id)
ORDER BY se.sample_date, s.site_code;")

core_arthropods[core_arthropods == ''] <- NA

core_arthropods <- core_arthropods %>% 
  mutate(
    sample_date = as.Date(sample_date),
    lt2mm = as.numeric(lt2mm),
    `_2_5mm` = as.numeric(`_2_5mm`),
    `_5_10mm` = as.numeric(`_5_10mm`),
    gt10mm = as.numeric(gt10mm),
    unsized = as.numeric(unsized),
    sampling_event_id = as.character(sampling_event_id)
  )

write_attributes(core_arthropods)
# write_factors(core_arthropods) # factors not present

core_arthropods_desc <- "Tabular file detailing key data from the CAP LTER's long-term monitoring of ground-dwelling arthropods. This data entity includes sample event details, taxonomic details and the number of organisms collected, sizes of organisms where noted (in the early stages of the project), and general notes and details regarding the gathering, identifiction, or enumeration of specimens."

core_arthropods_DT <- create_dataTable(dfname = core_arthropods,
                                       description = core_arthropods_desc,
                                       dateRangeField = 'sample_date')

```

```{r core_arthropods_sites}

core_arthropod_sites <- dbGetQuery(mysql, "
SELECT
  site_code,
  location,
  lat,
  `long`,
  gps_date,
  start_date,
  end_date,
  # num_traps,
  trap_arrange,
  comments
FROM lter10_arthropods_production.sites;")

core_arthropod_sites <- core_arthropod_sites %>% 
  # mutate(num_traps = as.numeric(num_traps)) %>% 
  mutate(
    start_date = as.Date(start_date),
    end_date = as.Date(end_date),
    gps_date = as.Date(gps_date)
  )

write_attributes(core_arthropod_sites) # write data frame attributes to a csv in current dir to edit metadata
# write_factors(core_arthropod_sites) # factors not present

core_arthropod_sites_desc <- "Tabular file providing detailed characteristics of arthropod sampling locations, including a general description of the location (typically nearest cross streets), latitude and longitude, sampling start and end dates (if applicable), arrangment of pitfall traps, and general comments regarding the sampling location."

core_arthropod_sites_DT <- create_dataTable(dfname = core_arthropod_sites,
                                            description = core_arthropod_sites_desc,
                                            dateRangeField = 'start_date')

```

```{r core_arthropods_kml}

library(sf)

core_arthropod_locations <- core_arthropod_sites %>%  
  filter(!is.na(lat))

core_arthropod_locations <- st_as_sf(x = core_arthropod_locations,
                                     coords = c("long", "lat"),
                                     crs = 4326)

# assign Name to the column that is the site identifier; drop geometry and empty
# columns to write attributes
core_arthropod_locations <- core_arthropod_locations %>% 
  mutate(Name = site_code) %>% 
  st_drop_geometry()

write_attributes(core_arthropod_locations)

# with attributes table written, repeat the process sans dropping the geometry
core_arthropod_locations <- core_arthropod_sites %>%  
  filter(!is.na(lat))

core_arthropod_locations <- st_as_sf(x = core_arthropod_locations,
                                     coords = c("long", "lat"),
                                     crs = 4326)

core_arthropod_locations <- core_arthropod_locations %>% 
  mutate(Name = site_code)

core_arthropod_locations_desc <- "Geospatial file (KML) detailing the locations of CAP LTER ground-dwelling arthropod sampling sites. Locations are approximately the center point of the set of traps at a given sampling site."

core_arthropod_locations_SV <- create_spatialVector(svname = core_arthropod_locations,
                                                    description = core_arthropod_locations_desc,
                                                    projectNaming = TRUE)

```

```{r title}
title <- 'Long-term monitoring of ground-dwelling arthropods in central Arizona-Phoenix, ongoing since 1998'
```

```{r abstract}

abstract <- set_TextType("abstract.md")
```

```{r people}

# see gioseml for examples of creating people resources from scratch

mysql_prod <- mysql_prod_connect()

nancyGrimm <- create_role(firstName = "n", lastName = "grimm", roleType = "creator")
danChilders <- create_role(firstName = "d", lastName = "childers", roleType = "creator")

creators <- list(nancyGrimm,
                 danChilders)

stevanEarl <- create_role(firstName = "s", lastName = "earl", roleType = "metadata")
sallyWittlinger <- create_role(firstName = "s", lastName = "wittling", roleType = "metadata")

metadataProvider <- list(stevanEarl,
                         sallyWittlinger)

# associated parties
markHostetler <- create_role(giosPersonId = 132)
nancyMcintyre <- create_role(giosPersonId = 150)
eyalShochat <- create_role(giosPersonId = 57)
stanFaeth <- create_role(giosPersonId = 184)

associatedParty <- list(stanFaeth,
                        markHostetler,
                        nancyMcintyre,
                        eyalShochat)

```

```{r keywords}

# CAP IRTs for reference (be sure to include these as appropriate):
# https://sustainability.asu.edu/caplter/research/

write_keywords()
keywords <- create_keywordSet('keywords.csv')

```

```{r methods}

methods <- set_methods("pitfall_trap_methods/pitfall_trapping_protocol_modified_from_v_sept2016.md")
```

```{r coverages}

begindate <- as.character(min(core_arthropods$sample_date))
enddate <- as.character(max(core_arthropods$sample_date))
geographicDescription <- "CAP LTER study area"
coverage <- set_coverage(begin = begindate,
                         end = enddate,
                         geographicDescription = geographicDescription,
                         west = -112.577118210717, east = -111.61479515834,
                         north = +33.8204250130355, south = +33.3028279145498)
```

New approach for taxonomy is to use taxonomyCleanr to build the
taxonomicCoverage. 

*Note* that the `taxa_map.csv` built with the `create_taxa_map()` function and
resolving taxonomic IDs (i.e., `resolve_comm_taxa()`) only needs to be run once
per version/session -- the taxonomicCoverage can be built as many times as
needed with `resolve_comm_taxa()` once the `taxa_map.csv` has been generated and
the taxonomic IDs resolved.

```{r taxonomyCleanr, eval=TRUE}

library(taxonomyCleanr)

my_path <- getwd() # taxonomyCleanr requires a path (to build the taxa_map)

arthropodTaxa <- core_arthropods %>% 
  mutate(
    display_name = str_replace(display_name, "\\(immature\\)", ""),
    display_name = str_replace(display_name, "\\(adult\\)", ""),
    display_name = str_replace(display_name, "\\(winged\\)", ""),
    display_name = str_replace(display_name, "\\(nymph\\)", ""),
    display_name = str_replace(display_name, "\\(larvae\\)", ""),
    display_name = str_replace(display_name, "\\(imm\\.\\)", ""),
    display_name = str_replace(display_name, "\\>10mm", ""),
    display_name = str_trim(display_name, side = c("both"))
  ) %>% 
  distinct(display_name) %>% 
  filter(!is.na(display_name))

# create or update map. A taxa_map.csv is the heart of taxonomyCleanr. This
# function will build the taxa_map.csv and put it in the path identified with
# my_path.
create_taxa_map(path = my_path, x = arthropodTaxa, col = "display_name") 

# resolve taxa by attempting to match the taxon's In this case, data.source 3 is
# ITIS (which, by the way, is the only authority taxonomyCleanr will allow for
# common names).
# resolve_comm_taxa(path = my_path, data.sources = 3) # in this case, 3 is ITIS
resolve_sci_taxa(path = my_path, data.sources = 3) # in this case, 3 is ITIS

# build the EML taxonomomic coverage
taxaCoverage <- make_taxonomicCoverage(path = my_path, write.file = FALSE)

# add taxonomic to other coverages
coverage$taxonomicCoverage <- taxaCoverage
```

```{r construct_dataset}

# from capeml package:
# address
# publisher
# contact
# rights
# distribution

# DATASET
dataset <- EML::eml$dataset(
  title = title,
  creator = creators,
  pubDate = pubDate,
  metadataProvider = metadataProvider,
  intellectualRights = capRights,
  abstract = abstract,
  keywordSet = keywords,
  coverage = coverage,
  contact = capContact, # cap contact
  publisher = capPublisher, # cap pub
  # contact = giosContact, # gios contact
  # publisher = giosPublisher, # gios pub
  methods = methods,
  project = capProject, # cap project
  distribution = create_distribution(packageIdent))

# add associatedParty if relevant
# dataset$associatedParty <- list() 
dataset$associatedParty <- associatedParty

```

```{r dataSet$dataTable}

# add dataTables if relevant

print(ls(pattern = "_DT"))

if (length(ls(pattern = "_DT")) > 0) {
  
  listOfDataTables <- lapply(ls(pattern = "_DT"), function(DT) { get(DT) } )
  
  dataset$dataTable  <- listOfDataTables  
  
}

# or add manually
# dataset$dataTable <- list(dataTableOne, dataTableTwo)

```

```{r dataSet$otherEntity}

# add other entities if relevant

print(ls(pattern = "_OE"))

if (length(ls(pattern = "_OE")) > 0) {
  
  listOfOtherEntities <- lapply(ls(pattern = "_OE"), function(OE) { get(OE) } )
  
  dataset$otherEntity <- listOfOtherEntities 
  
}

# or add manually
# dataset$otherEntity <- list(otherEntityOne, otherEntityTwo)

```

```{r dataSet$spatialVector}

# add spatial vectors if relevant

print(ls(pattern = "_SV"))

if (length(ls(pattern = "_SV")) > 0) {
  
  listOfSpatialVectors <- lapply(ls(pattern = "_SV"), function(SV) { get(SV) } )
  
  dataset$spatialVector  <- listOfSpatialVectors  
  
}

# or add manually
# dataset$spatialVector <- list(spatialVectorOne, spatialVectorTwo)

```

```{r literature cited, eval=FALSE}

# add literature cited if relevant
# dataset$literatureCited <- citationsAll 
```

```{r construct_eml, eval=TRUE}

if(exists('unitList')) {
  
  eml <- EML::eml$eml(
    access = lterAccess,
    dataset = dataset,
    additionalMetadata = unitList,
    packageId = packageIdent,
    system = "knb",
    scope = "system"
  )
  
} else {
  
  eml <- EML::eml$eml(
    access = lterAccess,
    dataset = dataset,
    packageId = packageIdent,
    system = "knb",
    scope = "system"
  )
}

```

```{r write_eml}

# write the eml to file
write_eml(eml, paste0(packageIdent, ".xml"))
```

```{r preview_data_file_to_upload}

# preview data set files that will be uploaded to S3
list.files(pattern = paste0(projectid, "_"))
```

```{r upload_data_S3}

# upload files to S3
lapply(list.files(pattern = paste0(projectid, "_")), dataToAmz)
```

```{r clean_up}

# remove data files
dataFilesToRemove <- dir(pattern = paste0(projectid, "_"))
file.remove(dataFilesToRemove)

# EML to S3
if(length(list.files(pattern = "*.xml")) == 1) {
  emlToAmz(list.files(pattern = "*.xml")) } else {
    print("more than one xml file found")
  }

# EML to cap-data-eml and remove file from project
tryCatch({
  
  if(length(list.files(pattern = "*.xml")) == 1) {
    file.copy(list.files(pattern = "*.xml"), "/home/srearl/localRepos/cap-metadata/cap-data-eml/")
    file.remove(list.files(pattern = "*.xml")) } else {
      print("more than one xml file found")
    }
},
warning = function(warn) {
  print(paste("WARNING: ", warn))
},
error = function(err) {
  print(paste("ERROR: ", err))
  
}) # close try catch
```
